---
title: "Database Space Usage"
author: "John Pertell"
date: "`r format(Sys.time(), '%a %b %d %Y %X')`"
output:
  html_document: 
    toc: TRUE
  
  
params:
  server1: "DCICHISQL1\\MISDB"
  server2: "DCICHISQL2\\MISDB"
  server3: "DCICORSQL3\\MISDB"
  server4: "DCICORSQL4\\MISDB"
  server5: "DCICORSQL8\\REGION5"
  db1: 'mis_db'
  db2: 'darwin_db'
  db3: 'dialysis_db'
---

# Introduction
This document shows the space usage over the last year on each production data server: 

  * `r params$server1`
  * `r params$server2`   
  * `r params$server3`   
  * `r params$server4`  
  * `r params$server5`    

It looks at the used space by each database file for the three largest database files:  

  * `r params$db1`
  * `r params$db2`
  * `r params$db3`
  
**NOTE: All sizes are in GB.**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RODBC)
library(dplyr)
library(ggplot2)
library(scales)
library(dygraphs)
library(xts)
library(quantmod)
library(tidyr)
library(zoo)

myConn <- odbcDriverConnect("driver={SQL Server};
                            server=(local);
                            database=DemoDB;
                            Trusted_Connection=yes")

db_filtered <- sqlQuery(myConn, "SELECT RTRIM(ServerName) AS ServerName, PollDate, DBName, DBLogicalFileName, DBFileSizeMB, 
                                        DBFileSizeMB - DBFreeSpaceMB AS DBUsedSpaceMB, DBFreeSpaceMB
                                 FROM DBInfo
                                 WHERE Servername IN ('DCICHISQL1\\MISDB', 'DCICHISQL2\\MISDB', 'DCICORSQL3\\MISDB', 'DCICORSQL4\\MISDB', 'DCICORSQL8\\REGION5')
                                 AND DBName IN ('mis_db', 'darwin_db', 'dialysis_db')
                                 AND PollDate >= DATEADD(year, -1, GETDATE())")

drive_filtered <- sqlQuery(myConn, "SELECT ServerName, Drive, DriveTotalSpaceMB, DriveFreeSpaceMB, DriveFreeSpacePct, PollDate
                                    FROM DSInfo
                                    WHERE Servername IN ('DCICHISQL1\\MISDB', 'DCICHISQL2\\MISDB', 'DCICORSQL3\\MISDB', 'DCICORSQL4\\MISDB', 'DCICORSQL8\\REGION5') 
                                    AND PollDate >= DATEADD(year, -1, GETDATE())")
close(myConn)

db_filtered$DBUsedSpaceGB <- (db_filtered$DBUsedSpaceMB / 1024) 
drive_filtered$DriveUsedSpaceGB <- (drive_filtered$DriveTotalSpaceMB - drive_filtered$DriveFreeSpaceMB) / 1024
```

# `r params$server1`  

## Drive Size Percent Free 

```{r echo = FALSE}
drive <- drive_filtered %>%
          filter(ServerName == params$server1) %>%
          group_by(Drive) %>%
          arrange(PollDate) %>%
          filter(row_number()==1 | row_number()==n()) %>% 
          select(PollDate, Drive, DriveFreeSpacePct) %>%
          spread(Drive, DriveFreeSpacePct)

as.data.frame(drive)
#ggplot(data = drive, aes(x = RunDate, y = pct_free, group = Drive )) + 
#  geom_line(aes(color = Drive)) 
``` 

## `r params$db1`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server1, DBName == params$db1) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB) 

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server1, DBName == params$db1) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName)) 

```  

## `r params$db2`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server1, DBName == params$db2) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB)  

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server1, DBName == params$db2) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName))  

```  

## `r params$db3`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server1, DBName == params$db3) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB)  

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server1, DBName == params$db3) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName))  

```

# `r params$server2`  

## Drive Size Percent Free 

```{r echo = FALSE}
drive <- drive_filtered %>%
          filter(ServerName == params$server2) %>%
          group_by(Drive) %>%
          arrange(PollDate) %>%
          filter(row_number()==1 | row_number()==n()) %>% 
          select(PollDate, Drive, DriveFreeSpacePct) %>%
          spread(Drive, DriveFreeSpacePct)

as.data.frame(drive)
#ggplot(data = drive, aes(x = RunDate, y = pct_free, group = Drive )) + 
#  geom_line(aes(color = Drive)) 
``` 

## `r params$db1`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server2, DBName == params$db1) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB)

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server2, DBName == params$db1) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName)) 

```  

## `r params$db2`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server2, DBName == params$db2) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB)

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server2, DBName == params$db2) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName)) 

```  

## `r params$db3`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server2, DBName == params$db3) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB)

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server2, DBName == params$db3) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName)) 

```

# `r params$server3`  

## Drive Size Percent Free 

```{r echo = FALSE}
drive <- drive_filtered %>%
          filter(ServerName == params$server3) %>%
          group_by(Drive) %>%
          arrange(PollDate) %>%
          filter(row_number()==1 | row_number()==n()) %>% 
          select(PollDate, Drive, DriveFreeSpacePct) %>%
          spread(Drive, DriveFreeSpacePct)

as.data.frame(drive)
#ggplot(data = drive, aes(x = RunDate, y = pct_free, group = Drive )) + 
#  geom_line(aes(color = Drive)) 
``` 

## `r params$db1`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server3, DBName == params$db1) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB) 

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server3, DBName == params$db1) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName))  

```  

## `r params$db2`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server3, DBName == params$db2) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB)  

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server3, DBName == params$db2) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName)) 

```  

## `r params$db3`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server3, DBName == params$db3) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB)  

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server3, DBName == params$db3) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName)) 

```

# `r params$server4`  

## Drive Size Percent Free 

```{r echo = FALSE}
drive <- drive_filtered %>%
          filter(ServerName == params$server4) %>%
          group_by(Drive) %>%
          arrange(PollDate) %>%
          filter(row_number()==1 | row_number()==n()) %>% 
          select(PollDate, Drive, DriveFreeSpacePct) %>%
          spread(Drive, DriveFreeSpacePct)

as.data.frame(drive)
#ggplot(data = drive, aes(x = RunDate, y = pct_free, group = Drive )) + 
#  geom_line(aes(color = Drive)) 
``` 

## `r params$db1`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server4, DBName == params$db1) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB)  

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server4, DBName == params$db1) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName))  

```  

## `r params$db2`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server4, DBName == params$db2) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB)  

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server4, DBName == params$db2) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName)) 

```  

## `r params$db3`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server4, DBName == params$db3) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB)

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server4, DBName == params$db3) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName)) 

```

# `r params$server5` 

## Drive Size Percent Free 

```{r echo = FALSE}
drive <- drive_filtered %>%
          filter(ServerName == params$server5) %>%
          group_by(Drive) %>%
          arrange(PollDate) %>%
          filter(row_number()==1 | row_number()==n()) %>% 
          select(PollDate, Drive, DriveFreeSpacePct) %>%
          spread(Drive, DriveFreeSpacePct)

as.data.frame(drive)
#ggplot(data = drive, aes(x = RunDate, y = pct_free, group = Drive )) + 
#  geom_line(aes(color = Drive)) 
``` 
## `r params$db1`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server5, DBName == params$db1) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB) 

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server5, DBName == params$db1) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName)) 

```  

## `r params$db2`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server5, DBName == params$db2) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB)  

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server5, DBName == params$db2) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName)) 

```  

## `r params$db3`
```{r echo = FALSE}
db_plot1 <- db_filtered %>%
  filter(ServerName == params$server5, DBName == params$db3) %>% 
  group_by(DBLogicalFileName) %>%
  arrange(PollDate) %>%
  filter(row_number()==1 | row_number()==n()) %>% 
  select(PollDate, DBLogicalFileName, DBUsedSpaceGB) %>%
  spread(DBLogicalFileName, DBUsedSpaceGB) 

db_plot2 <- db_filtered %>% 
              filter(ServerName == params$server5, DBName == params$db3) %>% 
              group_by(DBName) 

db_plot1 <- as.data.frame(db_plot1)
db_plot1

ggplot(data = db_plot2, aes(x = PollDate, y = DBUsedSpaceGB, group = DBLogicalFileName )) + 
  geom_line(aes(color = DBLogicalFileName)) 

```